Robust Bayesian Analysis of Causal Inference Problems

### 2022-10-28 JASA-template.tex
intro done

In causal inference, we are interested in estimating the causal
effect of independent variables on a dependent variable. Ideally,
randomised trials are the most efficient way to perform this task.
However, this is not very practical for several reasons; ethical 
concerns, design cost, population size, to name a few. This
leaves us with observational studies which are usually obtained
by means collecting data though surveys or record keeping. But this
can be problematic in the presence of confounders. That is when
the variables are associated with both the treatment and the outcome.
In such cases, we need to be extra cautious as otherwise it will
lead to unwanted bias in the treatment effect estimator \cite{rosenbaum83}.

Several works have been done in order to tackle the presence of
confounder variables. One such work in the topic was 
by \citet{Robins1986ANA} where the author used  a graphical
approach for the identification of the causal parameters.
\citet{rosenbaum1985} suggested the use of a link model to estimate
the propensity scores for all individuals. Later on several other
methods have been proposed based on propensity score matching.
A brief review on such methods can be found in \cite{winship99, stuart10}.

@@

In this paper we closely follow the approach of \citet{koch2020}, where
they proposed a bilevel spike and slab prior for causal effect 
estimation. They considered a data-driven adaptive approach to
reduce the variance of the causal estimate. Instead of that,
we suggest a robust Bayesian analysis based approach where instead
of using a single prior, we consider a set of priors \cite{BERGER1990303}. This is particularly interesting as in
many cases, causal effect estimation can be performed through a 
meta analysis and hence having a single prior can be problematic.
Moreover, for some problems we have to rely on very limited data to
perform our Bayesian analysis and choosing a data-driven prior
can lead to overfitting.

\begin{figure}
	\centering
	\begin{tikzpicture}[params/.style={circle, draw=green!60, fill=green!5, very thick, minimum size=7mm}]
	\node (1) at (0,0) {Treatment ($T$)};
	\node (2) [right = of 1] {Outcome ($Y$)};
	\node (4) [below = of 1] {Input ($X$)};

@@

Now, let $X\coloneqq$ $[X_1$, \dots, $X_n]^T$ 
denote the $p$-dimensional covariates present in the studies and $\beta \coloneqq (\beta_1$, \dots, $\beta_p)$ denotes the vector of regression
coefficients. Then we can define a linear model for the outcome
so that
\begin{equation}
    Y_i = \beta_{T} T_i + \beta_0 + X\beta + \epsilon_i
\end{equation}
where $\epsilon_i\sim \mathcal{N}(0, \sigma^2)$. Clearly, when
the underlying true outcome model is linear,
\begin{equation}
    \delta = \beta_{T}\quad\text{and hence } 
    \hat{\delta} - \delta = \hat{\beta}_T - \beta_T.
\end{equation}

@@

\section{Robust Bayesian Analysis}
\subsection{Hierarchical model}

@@

>>>>

In causal inference, we are interested in estimating the causal
effect of independent variables on a dependent variable. Ideally,
randomised trials are the most efficient way to perform this task.
However, this is not very practical for several reasons; ethical 
concerns, design cost, population size, to name a few. This
leaves us with observational studies which are usually obtained
by means collecting data though surveys or record keeping. But this
can be problematic in the presence of confounders. That is when
the variables are associated with both the treatment and the outcome.
In such cases, we need to be extra cautious as otherwise it will
lead to unwanted bias in the treatment effect estimator \cite{rosenbaum83}.
Several works have been done in order to tackle the presence of
confounder variables. One such work in the topic was 
by \citet{Robins1986ANA} where the author used  a graphical
approach for the identification of the causal parameters.
\citet{rosenbaum1985} suggested the use of a link model to estimate
the propensity scores for all individuals. Later on several other
methods have been proposed based on propensity score matching.
A brief review on such methods can be found in \cite{winship99, stuart10}.

@@

In this paper we take inspiration from the approach of \citet{koch2020}, where
they proposed a bilevel spike and slab prior for causal effect 
estimation. They considered a data-driven adaptive approach to
propose their prior which reduce the variance of the causal estimate. 
In our approach,
we suggest a sensitivity analysis based approach where instead
of using a single prior, we consider a set of priors \cite{BERGER1990303}. This is particularly interesting as in
many cases, causal effect estimation can be performed through a 
meta analysis and hence robust Bayesian analysis can be beneficial
\cite{raices_cruz22} under severe uncertainty.
Moreover, for some problems we have to rely on very limited data to
perform our Bayesian analysis and choosing a data-driven prior
can lead to overfitting. Therefore, as to propose our
robust Bayesian framework,
we consider a set of continuous spike and slab priors \cite{ishwaran2005}
for confounder selection and construct a Bayesian group LASSO
\cite{xu2015}. To perform the sensitivity analysis, we consider
a set of beta priors on the covariate selection probability
of the spike and slab priors. This sensitivity analysis allows us
to investigate the regression coefficients with respect to a different
level of sparsity of the model. Following that, we consider a post-hoc
variable selection method as suggested by \citet{hahn2015}. This
ensures that we can efficiently separate the confounder variables
with others and reduce the bias of treatment effect estimation.

The rest of the paper is organised as follows. In \cref{sec:causal}
we give a formal description of causal estimation problem in the
context of linear regression. \cref{sec:bayes} is focused on the
Bayesian analysis of causal inference problems, followed by the
motivation of a robust Bayesian analysis. In \cref{sec:sim}, we
provide our result of simulation studies under different 
scenarios and in data analysis with real data in \cref{sec:data:analysis}. Finally we discuss our findings and 
conclude this paper in \cref{sec:conc}.

\section{Causal Estimation}\label{sec:causal}

Now, let $X\coloneqq$ $[X_1$, \dots, $X_n]^T$ 
denote the $p$-dimensional covariates present in the studies and $\beta \coloneqq (\beta_1$, \dots, $\beta_p)$ denotes the vector of regression
coefficients. Then we can define a linear model for the outcome
so that
\begin{equation}
    Y_i = \beta_{T} T_i + \beta_0 + X\beta + \epsilon_i
\end{equation}
where $\epsilon_i\sim \mathcal{N}(0, \sigma^2)$. Clearly, when
the underlying true outcome model is linear,
\begin{equation}
    \delta = \beta_{T}\quad\text{and hence } 
    \hat{\delta} - \delta = \hat{\beta}_T - \beta_T.
\end{equation}
\begin{figure}
	\centering
	\begin{tikzpicture}[params/.style={circle, draw=green!60, fill=green!5, very thick, minimum size=7mm}]
	\node (1) at (0,0) {Treatment ($T$)};
	\node (2) [right = of 1] {Outcome ($Y$)};
	\node (4) [below = of 1] {Input ($X$)};

@@

@@

\section{Bayesian Hierarchical model}\label{sec:bayes}
\subsection{Hierarchical model}

@@

\section{Simulation Studies}\label{sec:sim}

\section{Data Analysis}\label{sec:data:analysis}

\section{Conclusion}\label{sec:conc}
